Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-05-04 doc/thesis/discussion.tex
Moar

The preferential attachment power \gls{alpha} had a very strong influence on
tree shape in the range of values we considered. Although the tree kernel was
the most effective classifier for \gls{alpha}, a tree balance statistic
performed nearly as well. This result was intuitive: high \gls{alpha} values
produce networks with few well-connected ``superspreader'' nodes which are
involved in a large number of transmissions, resulting in a highly unbalanced
ladder-like tree structure (\cref{fig:alphatrees}). The \gls{I} parameter,
representing the prevalence at the time of sampling, was also generally
estimable. The dynamics of the \gls{SI} model, and the coalescent
process~\autocite{kingman1982coalescent}, offer a potential explanation for
this result. \Cref{fig:sir} shows simulated trajectories of the \gls{SIR}
model. In the initial phase of the epidemic, when \gls{I} is small, each new
transmission results in potentially many new discordant edges, thus decreasing
the waiting time until the next transmission. Hence, there is an
early exponential growth phase, producing many short branches near the root of
the tree. As the epidemic gets closer to saturating the network, the number of
discordant edges decays, causing longer waiting times. The distribution of
coalescence times in the tree should therefore be informative about
\gls{I}~\autocite{volz2009phylodynamics}. This information is captured by the
tree kernel, and also by the \gls{nltt} statistic, which both performed quite
will in classifying \gls{I} (\cref{fig:rsquared}).

The number of nodes in the network, \gls{N}, exhibited the most variation in
terms of being estimable. There was almost no difference between trees
simulated under different \gls{N} values when the number of infected nodes
\gls{I} was very small. There is an intuitive explanation for this result,
namely that adding additional nodes does not change the edge density or overall
shape of a \gls{BA} network. This can be illustrated by imagining that we add a
small number of nodes to a network after the epidemic simulation has already
been completed. It is possible that none of these new nodes attains a
connection to any infected node. Thus, running the simulation again on the new,
larger network could produce the exact same transmission tree as before. On the
other hand, when \gls{I} is large relative to \gls{N}, the coalescent dynamics
discussed above also apply. That is, the waiting times until the next infection
increase, resulting in longer coalescence times toward the tips. The relative
accuracy of the \gls{nltt} in these situations coroborates this hypothesis.
The \gls{m} parameter, which controls the number of connections added to the
network per vertex, did not have a measurable impact on tree shape and was not
estimable with kernel-ABC. The exception to this was the value \gls{m} = 1,
which produces networks without cycles whose associated trees were more easily
distinguished. However, all the analyses presented here did not take the
absolute size of the transmission trees into account, as the branch lengths
were rescaled by their mean. Because higher \gls{m} values imply higher edge
density, an epidemic should spread more quickly for higher \gls{m} than lower
\gls{m} with the same per-edge transmission probability.  Hence, considering
the absolute height of the trees may improve our method's ability to
reconstruct \gls{m}.

@@

>>>>

The preferential attachment power \gls{alpha} had a very strong influence on
tree shape in the range of values we considered
(\cref{fig:alphatrees,fig:kpca}). Although the tree kernel was the most
effective classifier for \gls{alpha}, a Sackin's index of tree imbalance
performed nearly as well (\cref{fig:rsquared}). This result was intuitive: high
\gls{alpha} values produce networks with few well-connected ``superspreader''
nodes which are involved in a large number of transmissions, resulting in a
highly unbalanced ladder-like tree structure (\cref{fig:alphatrees}). There was
no observeable bias in the estimates of \gls{alpha} obtained with
\software{netabc}, however the variation in these estimates was higher for
$\alpha < 1$ than for $\alpha \geq 1$ (\cref{fig:abcptm2,tab:abchpd}). The
relationship between \gls{alpha} and the power law exponent \gls{gamma} may
explain this result (\cref{fig:gamma}). The \gls{gamma} values associated with
\gls{alpha} = 0 and \gls{alpha} = 0.5 are nearly identical (about 2.28 for
\gls{alpha} = 0 and 2.33 for \gls{alpha} = 0.5 with \gls{N} = 5000 and \gls{m}
= 2). In other words, the degree distributions of networks with \gls{alpha} $<
1$ are similar to each other, which may result in similarity of corresponding
transmission trees as well. 
%The accuracy of \gls{alpha} estimates also increased when the value of the
%\gls{m} parameter was higher (compare
%\cref{fig:abcptm2,fig:abcptm3,fig:abcptm4}).

\textbf{ART: THIS SECTION IS IN PROGRESS PLEASE DO NOT READ IT YET}

The \gls{I} parameter, representing the prevalence at the time of sampling, was
accurately classified by the tree kernel and the \gls{nltt}, but not by
Sackin's index (\cref{fig:rsquared}). When \gls{N} was held fixed, \gls{I} was
estimated accurately and without bias by grid search
\cref{fig:gridest,fig:gridI}. When the \gls{BA} model was fit with
\software{netabc}, the 95\% \gls{HPD} intervals for \gls{I} usually contained
the true value, but the \gls{MAP} estimates were biased slightly upward
(\cref{tab:abchpd}).

also generally estimable, although it was slightly over-estimated for both
cases we considered with kernel-\gls{ABC}.

The dynamics of the \gls{SI} model,
and the coalescent process~\autocite{kingman1982coalescent}, offer a potential
explanation for this result. In our simulations, we assumed that all discordant
edges shared the same transmission rate, so that the waiting time until the
next transmission in the entire network was always inversely proportional to
the number of discordant edges. Therefore, when there are a high number of
discordant edges,

In the initial phase of the epidemic, when \gls{I}
is small, each new transmission results in many new discordant edges.

thus
decreasing the waiting time until the next transmission. Hence,
there is an early exponential growth phase, producing many short branches near
the root of the tree. As the epidemic gets closer to saturating the network,
the number of discordant edges decays, causing longer waiting times. The
distribution of coalescence times in the tree should therefore be informative
about \gls{I}~\autocite{volz2009phylodynamics}. This information is captured by
the tree kernel, and also by the \gls{nltt} statistic, which both performed
quite well in classifying \gls{I} (\cref{fig:rsquared}).

\textbf{END OF IN PROGRESS}

The number of nodes in the network, \gls{N}, exhibited the most variation in
terms of its effect on tree shape. There was almost no difference between trees
simulated under different \gls{N} values when the number of infected nodes
\gls{I} was small. There is an intuitive explanation for this result, namely
that adding additional nodes does not change the edge density or overall shape
of a \gls{BA} network. This can be illustrated by imagining that we add a small
number of nodes to a network after the epidemic simulation has already been
completed. It is possible that none of these new nodes attains a connection to
any infected node. Thus, running the simulation again on the new, larger
network could produce the exact same transmission tree as before. On the other
hand, when \gls{I} is large relative to \gls{N}, the coalescent dynamics
discussed above also apply. That is, the waiting times until the next infection
increase, resulting in longer coalescence times toward the tips. The relative
accuracy of the \gls{nltt} in these situations coroborates this hypothesis.

%As mentioned above, \gls{I} and \gls{N} were both systematically over-estimated
%by our method, although the bias was more severe for \gls{N} than for \gls{I}.
%
%When \gls{N} was held fixed and \gls{I} was varied, grid search was able to
%recover \gls{I} without bias (\cref{fig:est}).

The \gls{m} parameter, which controls the number of connections added to the
network per vertex, did not have a measurable impact on tree shape and was not
estimable with kernel-ABC. The exception to this was the value \gls{m} = 1,
which produces networks without cycles whose associated trees were more easily
distinguished. However, all the analyses presented here did not take the
absolute size of the transmission trees into account, as the branch lengths
were rescaled by their mean. Because higher \gls{m} values imply higher edge
density, an epidemic should spread more quickly for higher \gls{m} than lower
\gls{m} with the same per-edge transmission probability.  Hence, considering
the absolute height of the trees may improve our method's ability to
reconstruct \gls{m}.

@@

The combination of method, model, and priors we employed did not produce
perfect estimates of any of the parameters. The estimates of \gls{alpha} were
the most accurate, although the variance of the estimates was high and the
confidence intervals were wide for \gls{alpha} $< 1$
(\cref{tab:abchpd,fig:abcptm2,fig:abcptm3,fig:abcptm4}). The estimates of
\gls{N} and \gls{I} were both biased, and the estimates of \gls{m} were largely
uninformative. Despite these issues, a major result of our our investigation is
that some contact network parameters have a measurable impact on tree shape
which can be used to perform statistical inference. Further refinements to
\software{netabc}, as well as the use of more sophisticated network models, may
improve the accuracy and precision of these estimates.

### 2016-05-04 doc/thesis/intro.tex
Moar

The kernel score of a pair of trees, denoted $K(T_1, T_2)$, is defined as a sum
over all pairs of nodes $(n_1, n_2)$, where $n_1$ is a node in $T_1$ and $n_2$
is a node in $T_2$. Folloting \textcite{poon2013mapping}, let $N(T)$ denote the
set of all nodes in $T$, $\nc(n)$ be the number of children of node $n$,
$c_n^{j}$ be the $j$th child of node $n$, and $l_n$ be the vector of branch
lengths connecting node $n$ to its descendants. The \defn{production rule} of
$n$ is its total number of children, and its number of leaf children. That is,
if two nodes have the same number of children and among these, the same number
of leaves, then they have the same production rule. Let $k_G(x, y)$ be a
Gaussian radial basis function of the vectors $x$ and $y$,
\[
  k_G(x, y) = \exp\left(-\frac{1}{2\sigma} \norm{x - y}_2^2\right),
\]
where $\norm{\cdot}_2$ is the Euclidian norm and $\sigma$ is a variance
parameter. The tree kernel is defined as 
\[
  K(T_1, T_2) = \sum_{n_1 \in N(T_1)} \sum{n_2 \in N(T_2)} \Delta (n_1, n_2),
\]
where
\[
  \Delta(n_1, n_2) =
  \begin{cases}
    \lambda & n_1 \text{ and } n_2 \text{ are leaves} \\
    \lambda k_G(l_{n_1}, l_{n_2}) \displaystyle\prod_{j=1}^{\nc(n_1)} \left(1 +
    \Delta(c_{n_1}^j, c_{n_2}^j) \right) & \begin{aligned} n_1 \text{ and } n_2 \text{ have the same} \\ \text{production rule} \end{aligned} \\
    0 & \text{otherwise}.

@@

Algorithms for performing \gls{ABC} fall into one of three categories:
rejection, \gls{MCMC}, and \gls{SMC}. To simplify the math, we shall restrict
the descriptions of these algorithms to the case of one simulated dataset per
parameter particle (the meaning of this will become clear shortly). The
extension to multiple datasets per particle is straightforward and will be
given at the end of the section. We use the variable $x$ to refer to the pair
$(\theta, z)$, so that the \gls{ABC} target distribution can be written
$\pi_\varepsilon(x \mid y)$.

>>>>

The kernel score of a pair of trees, denoted $K(T_1, T_2)$, is defined as a sum
over all pairs of nodes $(n_1, n_2)$, where $n_1$ is a node in $T_1$ and $n_2$
is a node in $T_2$. Following \textcite{poon2013mapping}, let $N(T)$ denote the
set of all nodes in $T$, $\nc(n)$ be the number of children of node $n$,
$c_n^{j}$ be the $j$th child of node $n$, and $l_n$ be the vector of branch
lengths connecting node $n$ to its descendants. The \defn{production rule} of
$n$ is its total number of children, and its number of leaf children. That is,
if two nodes have the same number of children and among these, the same number
of leaves, then they have the same production rule. Let $k_G(x, y)$ be a
Gaussian radial basis function of the vectors $x$ and $y$,
\[
  k_G(x, y) = \exp\left(-\frac{1}{2\sigma} \norm{x - y}_2^2\right),
\]
where $\norm{\cdot}_2$ is the Euclidian norm and $\sigma$ is a variance
parameter. The tree kernel is defined as 
\[
  K(T_1, T_2) = \sum_{n_1 \in N(T_1)} \sum{n_2 \in N(T_2)} \Delta (n_1, n_2),
\]
where
\[
  \Delta(n_1, n_2) =
  \begin{cases}
    \lambda & n_1 \text{ and } n_2 \text{ are leaves} \\
    \lambda k_G(l_{n_1}, l_{n_2}) \displaystyle\prod_{j=1}^{\nc(n_1)} \left(1 +
    \Delta(c_{n_1}^j, c_{n_2}^j) \right) & \begin{aligned} n_1 \text{ and } n_2 \text{ have the same} \\ \text{production rule} \end{aligned} \\
    0 & \text{otherwise}.

@@

Algorithms for performing \gls{ABC} fall into one of three categories:
rejection, \gls{MCMC}, and \gls{SMC}~\autocite{marin2012approximate}. To
simplify the math, we shall restrict the descriptions of these algorithms to
the case of one simulated dataset per parameter particle (the meaning of this
will become clear shortly). The extension to multiple datasets per particle is
straightforward and will be given at the end of the section. We use the
variable $x$ to refer to the pair $(\theta, z)$, so that the \gls{ABC} target
distribution can be written $\pi_\varepsilon(x \mid y)$.

### 2016-05-04 doc/thesis/methods.tex
Moar

We investigated four parameters related to the \gls{BA} model, denoted
\gls{N}, \gls{m}, \gls{alpha}, \gls{I}. The first three of these are parameters
of the model itself, while \gls{I} is related to the simulation of transmission
trees over the network. However, we will refer to all four as \gls{BA}
parameters. \gls{N} denotes the total number of nodes in the network, or
equivalently, susceptible individuals in the population. When a node is added
to the network, \gls{m} new undirected edges are added incident to it, and are
attached to existing nodes of degree $k$ with probability proportional to
$k^\alpha + 1$ (\cref{subsec:pa}). To simulate transmission trees over a
\gls{BA} network, we allowed an epidemic to spread until \gls{I} nodes were
infected, and sampled a transmission tree at that time. We assumed that all
contacts had symmetric transmission risk, which was implemented by replacing
each undirected edge in the network with two directed edges (one in each
direction).

Nodes in our networks followed simple \gls{SI} dynamics, meaning that they
became infected at a rate proportional to their number of infected neighbours,
and never recovered.  We did not consider the time scale of the transmission
trees in these simulations, only their shape. Therefore, the transmission rate
along each edge in the network was set to 1, the removal rate of each node was
set to 0, and all transmission trees' branch lengths were scaled by their mean.
For all analyses, the transmission trees' branch lengths were scaled by
dividing by their mean. We used the \textit{igraph} library's implementation of
the BA model~\autocite{csardi2006igraph} to generate the graphs. The analyses
were run on Westgrid (\url{https://www.westgrid.ca/}) and a local computer
cluster.

\subsubsection*{Kernel classifiers}
\label{subsec:kernel}

The experiments presented here involved a large number of variables which were
varied combinatorially. For ease of exposition, we will describe a single
experiment first, then enumerate the values of all variables for which the
experiment was repeated. The parameters of the tree kernel, $\lambda$ and
$\sigma$ (\cref{subsec:treeshape}) will be referred to as
\defn{meta-parameters} to distinguish them from the parameters of the \gls{BA}
model. With the exception of our own programs, all analyses were done in
\software{R}, and all packages listed below are \software{R} packages.

@@

As in the previous section, we will begin by describing a single experiment,
and then list the variables for which similar experiments were performed. We
varied \gls{alpha} along a narrowly spaced grid of values: 0, 0.01, \ldots, 2.
For each value, fifteen networks were generated with \software{igraph}, and
transmission trees were simulated over each using \software{nettree}. These
trees will be referred to as ``grid trees'', and their associated values ``grid
values''. Next, one further test tree was simulated with the test value
\gls{alpha} = 0. Both the grid trees and the test tree had 500 tips, and were
simulated with the other \gls{BA} parameters set to $N$ = 5000, $m$ = 2, and
$I$ = 1000. The test tree was compared to each of the grid trees using the tree
kernel, with the meta-parameters set to $\lambda = 0.3$ and $\sigma = 4$, using
the \software{treekernel} binary. The median kernel score was calculated for
each grid value, and the scores were normalized such that the area under the
curve was equal to 1. The grid value with the highest median kernel score was
taken as the point estimate for the test value, and a 95\% credible interval
was obtained using the \software{hpd} function in the \software{TeachingDemos}
package.

Each experiment of the type just described was repeated ten times with the same
test value. Similar experiments were performed for each of the four \gls{BA}
parameters, with several test values and trees of varying sizes. The variables
are listed in \cref{tab:gridexpt}. A graphical schematic of the grid search
experiments is shown in \cref{fig:gridexpt}.

@@

%\subsubsection*{Characterization of power-law exponent in Barab\'asi-Albert networks}
%
%Most studies of social network or transmission network
%parameters~\autocite[e.g.][]{liljeros2001web, jones2003assessment,
%schneeberger2004scale, brown2011transmission} report the coefficient
%\gls{gamma} of the power law degree distribution. To make our results
%comparable to previous work, we used simulated networks to investigate the
%relationship between the \gls{BA} model parameters and \gls{gamma}. A network
%was simulated for each combination of parameters listed in
%\cref{tab:gammaexpt}. A power law distribution was fitted to the degree
%distribution of each simulated network using the \software{fit\_power\_law}
%function in \software{igraph} with the `R.mle' implementation. We fitted a
%\gls{GLM} with Gamma-distributed errors and a log link function to the observed
%distribution of \gls{gamma} values, with \gls{alpha}, \gls{m}, \gls{N}, and all
%possible interaction terms as predictors.

\begin{table}
  \centering
  \input{\tablepath/gamma-expt}
  \caption[\gls{BA} parameters used as input \gls{GLM} predicting $\gamma$]
  {
    \gls{BA} model parameters used as input to \gls{GLM} predicting power law
    exponent $\gamma$. One network was simulated with each combination of
    parameters, and $\gamma$ was calculated for each network. A \gls{GLM} with
    Gamma-distributed errors and a log link function was fit to the $\gamma$
    values with all parameters and interaction terms as predictors.
  }
  \label{tab:gammaexpt}
\end{table}

>>>>

We investigated four parameters related to the \gls{BA} model, denoted
\gls{N}, \gls{m}, \gls{alpha}, \gls{I}. The first three of these are parameters
of the model itself, while \gls{I} is related to the simulation of transmission
trees over the network. However, we will refer to all four as \gls{BA}
parameters. \gls{N} denotes the total number of nodes in the network, or
equivalently, susceptible individuals in the population. When a node is added
to the network, \gls{m} new undirected edges are added incident to it, and are
attached to existing nodes of degree $k$ with probability proportional to
$k^\alpha + 1$ (\cref{subsec:pa}). To simulate transmission trees over a
\gls{BA} network, we allowed an epidemic to spread until \gls{I} nodes were
infected, and sampled a transmission tree at that time. The \gls{alpha}
parameter is unitless, while \gls{m} has units of edges or connections per
vertex, and \gls{N} and \gls{I} both have units of nodes or individuals.

We assumed that all contacts had symmetric transmission risk, which was
implemented by replacing each undirected edge in the network with two directed
edges (one in each direction). Nodes in our networks followed simple \gls{SI}
dynamics, meaning that they became infected at a rate proportional to their
number of infected neighbours, and never recovered. We did not consider the
time scale of the transmission trees in these simulations, only their shape.
Therefore, the transmission rate along each edge in the network was set to 1,
the removal rate of each node was set to 0, and all transmission trees' branch
lengths were scaled by their mean. For all analyses, the transmission trees'
branch lengths were scaled by dividing by their mean. We used the
\textit{igraph} library's implementation of the BA
model~\autocite{csardi2006igraph} to generate the graphs. The analyses were run
on Westgrid (\url{https://www.westgrid.ca/}) and a local computer cluster. With
the exception of our own programs, all analyses were done in \software{R}, and
all packages listed below are \software{R} packages.

\subsubsection*{Classifiers for BA model parameters based on tree shape}
\label{subsec:kernel}

The experiments presented here involved a large number of variables which were
varied combinatorially. For ease of exposition, we will describe a single
experiment first, then enumerate the values of all variables for which the
experiment was repeated. The parameters of the tree kernel, $\lambda$ and
$\sigma$ (\cref{subsec:treeshape}) will be referred to as
\defn{meta-parameters} to distinguish them from the parameters of the \gls{BA}
model.

@@

As in the previous section, we will begin by describing a single experiment,
and then list the variables for which similar experiments were performed. We
varied \gls{alpha} along a narrowly spaced grid of values: 0, 0.01, \ldots, 2.
For each value, fifteen networks were generated with \software{igraph}, and
transmission trees were simulated over each using \software{nettree}. These
trees will be referred to as ``grid trees'', and their associated values ``grid
values''. Next, one further test tree was simulated with the test value
\gls{alpha} = 0. Both the grid trees and the test tree had 500 tips, and were
simulated with the other \gls{BA} parameters set to $N$ = 5000, $m$ = 2, and
$I$ = 1000. The test tree was compared to each of the grid trees using the tree
kernel, with the meta-parameters set to $\lambda = 0.3$ and $\sigma = 4$, using
the \software{treekernel} binary. The median kernel score was calculated for
each grid value, and the scores were normalized such that the area under the
curve was equal to 1. The grid value with the highest median kernel score was
taken as the point estimate for the test value, and a 95\% credible interval
was obtained using the \software{hpd} function in the \software{TeachingDemos}
package~\autocite{snow2013teachingdemos}.

Each experiment of the type just described was repeated ten times with the same
test value. Similar experiments were performed for each of the four \gls{BA}
parameters, with several test values and trees of varying sizes. The variables
are listed in \cref{tab:gridexpt}. A graphical schematic of the grid search
experiments is shown in \cref{fig:gridexpt}. 
%Spearman's correlation was used to test whether the number of tips in the tree
%was correlated with the acccuracy of the point estimates. One-way \gls{ANOVA}
%was used to test whether the accuracy of the point estimates were significantly
%higher for any parameter values. If so, the distribution of errors was examined
%to choose a suitable post hoc test. If there appeared to be a correlation
%between the parameter value and the accuracy of the estimate, Spearman's
%correlation was calculated. If there were one or more particular values which
%appeared to have lower error, we used the Wilcoxon rank-sum test.

@@

%\subsubsection*{Characterization of power-law exponent in Barab\'asi-Albert networks}
%
%Most studies of social network or transmission network
%parameters~\autocite[e.g.][]{liljeros2001web, jones2003assessment,
%schneeberger2004scale, brown2011transmission} report the coefficient
%\gls{gamma} of the power law degree distribution. To make our results
%comparable to previous work, we used simulated networks to investigate the
%relationship between the \gls{BA} model parameters and \gls{gamma}. A network
%was simulated for each combination of parameters listed in
%\cref{tab:gammaexpt}. A power law distribution was fitted to the degree
%distribution of each simulated network using the \software{fit\_power\_law}
%function in \software{igraph} with the `R.mle' implementation. We fitted a
%\gls{GLM} with Gamma-distributed errors and a log link function to the observed
%distribution of \gls{gamma} values, with \gls{alpha}, \gls{m}, \gls{N}, and all
%possible interaction terms as predictors.
%
%\begin{table}
%  \centering
%  \input{\tablepath/gamma-expt}
%  \caption[\gls{BA} parameters used as input \gls{GLM} predicting $\gamma$]
%  {
%    \gls{BA} model parameters used as input to \gls{GLM} predicting power law
%    exponent $\gamma$. One network was simulated with each combination of
%    parameters, and $\gamma$ was calculated for each network. A \gls{GLM} with
%    Gamma-distributed errors and a log link function was fit to the $\gamma$
%    values with all parameters and interaction terms as predictors.
%  }
%  \label{tab:gammaexpt}
%\end{table}

%To evaluate the effects of the true parameter values on the accuracy of the
%\gls{MAP} estimates, we analysed each parameter individually using a \gls{GLM}
%(four \glspl{GLM} total). The response variable was the error of the point
%estimate, and the predictor variables were the true values of \gls{alpha},
%\gls{I}, and \gls{m}. We did not test for differences across true values of
%\gls{N}, because \gls{N} was not varied in these simulations. The distribution
%family and link function for each \gls{GLM} were chosen by examination of
%residual plots and \gls{AIC} of the model fits. For \gls{alpha}, \gls{I}, and
%\gls{N}, we used the Gamma distribution with the inverse link function. For
%\gls{m}, we used the Poisson distribution with the log link function. To
%allow for the possibility that the relationships between parameters and error
%rates were nonlinear, the true parameter values were treated as discrete rather
%than continuous predictors. When significant associations were found by the
%\glspl{GLM}, additional post hoc tests were performed based on the observed
%relationship between the error rate and true parameter value in question. If
%the relationship appeared to be monotonic, Spearman's correlation was
%calculated. Otherwise, if there were particular values which appeared to have a
%higher error rate than the others, the Wilcoxon rank-sum test was used to check
%for a significant difference between the two groups.

### 2016-05-04 doc/thesis/preface.tex
Moar

The initial idea to use kernel approximate Bayesian computation to infer
contact network model parameters was Dr. Poon's. The tree kernel was originally
developed by Dr. Poon, but the version used here was implemented by me to
improve computational efficiency. The idea to apply sequential Monte Carlo is
credited to Dr. Alexandre Bouchard-C\^ot\'e. Dr. Sarah Otto suggested the
experiments involving a network with a heterogeneous $\alpha$ parameter and
peer-driven sampling. Dr. Richard H. Liang provided guidance in the development
of the Gillespie simulation algorithm. The \software{netabc} program, and all
supplementary analysis programs, were written by me.

A version of chapter 2 has been submitted for publication in Virus Evolution
with the title ``Reconstructing network parameters from viral phylogenies.'' An
oral presentation entitled ``Phylodynamic inference of contact network
parameters with kernel-ABC'' was given based on chapter 2 at the 23rd HIV
Dynamics and Evolution meeting on April 25, in Woods Hole, Massachussets, USA.
A poster based on chapter 2 entitled ``Likelihood-free estimation of contact
network parameters from viral phylogenies'' was presented at the Intelligent
Systems for Molecular Biology meeting in Orlando, Florida, USA.

>>>>

The initial idea to use kernel approximate Bayesian computation to infer
contact network model parameters was Dr. Poon's. The tree kernel was originally
developed by Dr. Poon, but the version used here was implemented by me to
improve computational efficiency. The idea to apply sequential Monte Carlo was
mine, but Dr. Alexandre Bouchard-C\^ot\'e made me aware of the adaptive version
used in this work. Dr. Sarah Otto suggested the experiments involving a network
with a heterogeneous $\alpha$ parameter and peer-driven sampling. Dr. Richard
H. Liang provided guidance in the development of the Gillespie simulation
algorithm and statistical advice. The \software{netabc} program, and all
supplementary analysis programs, were written by me.

A version of chapter 2 has been submitted for publication in Virus Evolution
with the title ``Reconstructing network parameters from viral phylogenies.'' An
oral presentation entitled ``Phylodynamic inference of contact network
parameters with kernel-ABC'' was given based on chapter 2 to the 23rd HIV
Dynamics and Evolution meeting on April 25, in Woods Hole, Massachussets, USA
(the presentation was delivered remotely). A poster based on chapter 2 entitled
``Likelihood-free estimation of contact network parameters from viral
phylogenies'' was presented at the Intelligent Systems for Molecular Biology
meeting in Orlando, Florida, USA.

Use of the BC data is in accordance with an ethics application that was
reviewed and approved by the UBC/Providence Health Care Research Ethics Board
(H07-02559).
